# Simple description

- Describe labor market after 2019.

## Employment rate

- Report $e_{g,m,y} = \frac{Employment_{g,m,y}}{Population_{g,m,y}}$, where $Employment_{g,m,y}$ and $Population_{g,m,y}$ are numbers of employment and population over 15 years old in month $m$, year $y$ and gender group $g$, respectively.

```{r}
raw  %>%
  filter(year >= 2019) %>%
  mutate(y = employment_rate) %>%
  ggplot(aes(x = date,
             y = y,
             color = sex)
         ) +
  geom_line() +
  geom_point() +
  theme_bw() +
  ylab("employment rate")
```

## Year-to-year difference of employment rate

- Report change of employment rate $\tilde e_{g,m,y}=e_{g,m,y}-e_{g,m,y-1}$

```{r}
raw %>%
  arrange(sex,
          quaterly,
          year) %>%
  group_by(sex,
           quaterly) %>%
  mutate(y = employment_rate - lag(employment_rate)) %>%
  ungroup %>%
  filter(year >= 2019) %>%
  ggplot(aes(x = date,
             y = y,
             color = sex)
         ) +
  geom_line() +
  geom_point() +
  geom_hline(yintercept = 0) +
  ylab("") +
  xlab("") +
  theme_bw() +
  theme(legend.position = "bottom")
```


## Gender gap

- Report change of employment rate $\tilde e_{male,m,y} - \tilde e_{female,m,y}$

```{r}
raw %>%
  arrange(sex,
          quaterly,
          year) %>%
  group_by(sex,
           quaterly) %>%
  mutate(y = employment_rate - lag(employment_rate)) %>%
  ungroup %>%
  filter(year >= 1969) %>%
  arrange(date,
          sex) %>%
  group_by(date) %>%
  mutate(y = y - lag(y)) %>%
  ungroup %>%
  filter(sex == "male") %>%
  filter(year >= 2019) %>%
  ggplot(aes(x = date,
             y = y)
         ) +
  geom_line() +
  geom_point() +
  geom_hline(yintercept = 0) +
  ylab("") +
  xlab("") +
  theme_bw()
```


